Towards a miniKanren with fair search strategies

### 2019-05-21 Paper/paper.tex
These are in Paper

\addeditor{MVC}
\addeditor{LKC}

@@

\title{miniKanren with fair search strategies}
\author{Kuang-Chen Lu}
\affiliation{Indiana University}
\author{Weixi Ma}
\affiliation{Indiana University}
\author{Daniel P. Friedman}
\affiliation{Indiana University}

@@

\newtheorem{defn}{Definition}[section]

@@

We describe fairness levels in disjunction implementation and conjunction 
implementation. Specifically, a disjunction implementation can be fair, 
almost-fair, or unfair. And a conjunction implementation can be fair or unfair. 
We compare the fairness level of four search strategies: interleaving 
depth-first search (DFSi), balanced interleaving depth-first search (DFSbi), 
fair depth-first search (DFSf), and breadth-first search (BFS). DFSi is one of 
the best-understood search strategies in miniKanren community. DFSbi and DFSf 
are new. And we present a new implementation of BFS. We demonstrate by a 
quantitative evaluation that DFSbi and DFSf are competitive alternatives to 
DFSi, and that BFS is less practical.

@@

miniKanren is a family of relational programming languages.
\citet{friedman_reasoned_2018} introduce miniKanren and its implementation
in \emph{The Reasoned Schemer, 2nd Ed} (TR2). miniKanren programs, especially
relational interpreters, have been proven to be useful in solving many problems 
by \citet{byrd2017unified}.

\NOTE[LKC]{Shall we delete the sentence about relational interpreters? It is 
misleading: it is emphasizing relational interpreters, but we say nothing 
further about them later.}

% \conde{} is the flagship disjunction operator in miniKanren.
A subtlety arises 
when a \conde{} contains many clauses: not every clause has an 
equal chance to contribute to the result. As an example, consider the following 
relation \texttt{repeato} and its invocation.

@@

The miniKanren in TR2 implements interleaving DFS (\DFSi), the cause of this 
unexpected result. With this search strategy, each clause takes half 
of its received computational resources and pass the other half to its 
following clauses, except for the last clause that takes all resources it 
receives. In the example above, the \texttt{a} clause takes half of all 
recourses. And the \texttt{b} clause takes a quarter. Thus \texttt{c} and 
\texttt{d} barely contribute to the result.

@@

\DFSi{} is sometimes powerful for an expert. By carefully organizing the order 
of \conde{} clauses, a miniKanren program can explore more ``interesting'' 
clauses than those uninteresting ones, and thus use computational resources 
efficiently. A little miniKanrener, however, may beg to differ--understanding
implementation details and fiddling with clauses order is not the first
priority of a beginner.

@@

The specification that every clause in the same \conde{} is given equal 
``search priority'' is called fair \disj{}. And search strategies with 
almost-fair \disj{} give every clause similar priority. 
Fair \conj{}, a related concept, is more complicated. We defer it to the next 
section.

To summarize our contribution, we
\begin{itemize}
	\item propose and implement balanced interleaving depth-first search 
	(\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement fair depth-first search (\DFSf{}), a 
	new search strategy with fair \disj{}.
	\item implement in a new way breath-first search (BFS), a search strategy 
	with fair \disj{} and fair \conj{}. And we prove formally that our BFS 
implementation is equivalent to the one by \citet{seres1999algebra}. Our code 
runs faster in all benchmarks and is simpler. 
\end{itemize}

@@

Under the hood, the \conde{} here is allocating computational resource to 
four trivially different search space. The unfair \disj{} in 
\DFSi{} allocates many more resources to the first search space. On the 
contrary, fair \disj{} would allocate resources evenly to each search space.

@@

We end this subsection with precise definitions of all levels of 
\disj{} fairness. Our definition of \emph{fair} \disj{} is slightly generalize 
from the one by \citet{seres1999algebra}. Their definition is for binary 
disjunction. We generalize it to multi-arity one.

\begin{defn}[fair \disj{}]
A \disj{} is fair if and only if it allocates computational resource evenly to 
search spaces produced by goals in the same disjunction (i.e. clauses in 
the same \conde).
\end{defn}

\begin{defn}[almost-fair \disj{}]
A \disj{} is almost-fair if and only if it allocates computational resource so 
evenly to 
search spaces produced by goals in the same disjunction that the maximal ratio 
of resources is bounded by a constant.
\end{defn}

@@

A more interesting situation is when the first conjunct produces infinite many 
states. Consider the following example, a naive specification of fair \conj{} 
might require search strategies to produce all sorts of singleton lists, but no 
longer ones, which makes the strategies incomplete. A search strategy is 
\emph{complete} if and only if it can find out all the answers within finite 
time, otherwise, it is \emph{incomplete}.

@@

BFS puts states of the same cost in the same bag, and allocate resources 
carefully among search spaces related to different bags such that answers are 
produced in increasing order of cost. The \emph{cost} of a answer is its depth 
in the search tree (i.e. the number of calls to relational definitions required 
to find them) \citep{seres1999algebra}. In the following example, every answer 
is a list of list of symbols, where inner lists are the same. Here the 
cost of each answer is equal to the length of its inner lists plus the length 
of its outer list.

@@

\begin{defn}[fair \conj{}]
A \conj{} is fair if and only if it allocates computational resource evenly to 
search spaces produced from states in the same bag. A bag is a finite 
collection of state. And search strategies with fair \conj{} should represent 
search spaces with possibly infinite collection of bags. 
\end{defn}

@@

Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} shows parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: variables bound to 
states are named \texttt{s}; variables bound to goals are named \texttt{g} 
(possibly with subscript); variables bound to search spaces have names ending 
with $^\infty$. Fig.~\ref{DFSi-0} shows the implementation of disjunction. The 
first function, \texttt{disj2}, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say disjunctive operator is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of conjunction. The first function, \texttt{conj2}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The latter process is done with a 
helper function. \texttt{append-map$^\infty$} applies its input goal to states 
in its input search spaces and compose the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say conjunctive operator is also right-associative.

@@

Balanced interleaving DFS (DFS$_\textrm{bi}$) has almost-fair \disj{} and unfair 
\conj{}. The implementation of DFS$_\textrm{bi}$ differs from 
DFS$_\textrm{i}$'s in the \disj{} macro. We list the new \disj{} with its 
helper in Fig.~\ref{balanced-disj}. When there are one or more disjunctive 
goals, \disj{} builds a balanced binary tree whose leaves are the goals and 
whose nodes are \texttt{disj2}s, hence the name of this search strategy. 
The new helper, \texttt{disj+}, takes two additional `arguments'. They 
accumulate goals to be put in the left and right subtrees. The first clause 
handles the case where there is only one goal. In this case, the tree is the 
goal itself. When there are more goals, we partition the list of goals 
into two sublists of roughly equal lengths and recur on the two sublists. Goals 
are moved to the accumulators in the fourth and last clause. As we are moving 
two goals each time, there are two base cases: (1) no goal remains; (2) one 
goal remains. These two base cases are handled in the second clause and the 
third clause respectively. In contrast, the \disj{} in \DFSi{} constructs the 
binary tree in a particularly unbalanced form.

@@

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\texttt{disj2} (Fig.~\ref{fDFS}). \texttt{disj2} is changed to call a new and 
fair version of \texttt{append$^\infty$}. \texttt{append$^\infty_{fair}$} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, set to 
\texttt{\#{}t}, which indicates that 
\texttt{s$^\infty$} and \texttt{t$^\infty$} haven't been swapped. The swapping 
happens at 
the third \texttt{cond} clause in the helper, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \texttt{s$^\infty$} and 
\texttt{t$^\infty$} are ended with a thunk and they have been swapped. In this 
case, a new thunk is constructed. Here changing the order of 
\texttt{t$^\infty$} and \texttt{s$^\infty$} won't hurt the fairness (though it 
will change the order of answers). We swap them back so that answers are 
produced in a more natural order.

@@

BFS has both fair \disj{} and fair \conj{}. Our implementation is based on 
DFS$_\textrm{f}$ (not DFS$_\textrm{i}$). To implement \BFS{} based on \DFSf{}, 
we need \texttt{append-map$^\infty_{fair}$} in addition to 
\texttt{append$^\infty_{fair}$}. 
The only difference between \texttt{append-map$^\infty$} and 
\texttt{append-map$^\infty_{fair}$} is that the latter calls 
\texttt{append$^\infty_{fair}$} instead of \texttt{append$^\infty$}.

The implementation can be improved in two ways. First, as mentioned in 
section 2.2, BFS puts answers in bags and answers of the same cost are in the 
same bag. In this implementation, however, it is unclear where this information 
is recorded. Second, \texttt{append$^\infty_{fair}$} is extravagant in memory 
usage. It 
makes $O(n+m)$ new \texttt{cons} cells every time, where $n$ and $m$ are the 
``length''s of input search spaces. We address these issues in the first 
subsection.

Both our BFS and Seres's BFS \citet{seres1999algebra} produce answers in 
increasing order of cost. So it is interesting to see if they are equivalent. 
We prove so in Coq. The details are in the second subsection.

@@

As mentioned in section 2.2, BFS puts answers in bags and answers of the 
same cost are in the same bag. The cost
information is recorded subtly -- the \texttt{car}s of a search space have cost 
0 (i.e. they are in the same bag), and the costs of answers in thunk are 
computed recursively then increased by one. It is even more subtle that
\texttt{append$^\infty_{fair}$} and the \texttt{append-map$^\infty_{fair}$} 
respects 
the cost 
information. We make these facts more obvious by changing the type of search 
space, modifying related function definitions, and introducing a few more 
functions.

@@

Functions related to the pure subset are listed in Fig.~\ref{BFS-opt} (the 
others in Fig.~\ref{BFS-opt-cont}). They are compared with 
\citeauthor{seres1999algebra}'s implementation in our proof. The first three 
functions in Fig.~\ref{BFS-opt} are search space constructors. \texttt{none} 
makes an empty search space; \texttt{unit} makes a space from one answer; and 
\texttt{step} makes a space from a thunk. The remaining functions do the same 
thing as before.

Luckily, the change in \texttt{append$^\infty_{fair}$} also fixes the miserable 
space 
extravagance -- the use of \texttt{append} helps us to reuse the first bag of 
\texttt{t$^\infty$}.

\citet{kiselyov2005backtracking} has shown that a \emph{MonadPlus} hides in 
implementations of logic programming system. Our BFS implementation is not an 
exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, 
but takes arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, 
and \texttt{mplus} respectively.

@@

\subsection{comparison with the BFS of \citet{seres1999algebra}}

In this section, we compare the pure subset of our optimized BFS with the BFS 
found in \citet{seres1999algebra}. We focus on the pure subset because 
their system is pure. Their system represents search spaces with streams of 
lists of answers, where each list is a bag.

To compare efficiency, we translate their Haskell code into Racket (See 
supplements for the translated code). The translation is direct 
due to the similarity in both logic programming systems and search space 
representations. The translated code is longer and slower than our BFS 
implementation. Details about 
difference in efficiency are in section 6.

We prove in Coq that the two BFSs are equivalent, i.e. \texttt{(run n g)} 
produces the same result (See supplements for the formal proof).

@@

\begin{table}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline 
		benchmark & size & DFS$_\textrm{i}$ & DFS$_\textrm{bi}$ & DFS$_\textrm{f}$ & optimized BFS & Silvija's BFS  
		\\
		\hline
		very-recursiveo & 100000 &  579 &  793 & 2131 & 1438 & 3617 \\
		& 200000 & 1283 & 1610 & 3602 & 2803 & 4212 \\
		& 300000 & 2160 & 2836 &    - & 6137 &    - \\
		\hline 
		appendo  & 100 &  31 &  41 &  42 &  31 &  68 \\ 
		& 200 & 224 & 222 & 221 & 226 & 218 \\ 
		& 300 & 617 & 634 & 593 & 631 & 622 \\ 
		\hline 
		reverso & 10 &   5 &   3 &   3 &     38 &     85 \\ 
		& 20 & 107 &  98 &  51 &   4862 &   5844 \\
		& 30 & 446 & 442 & 485 & 123288 & 132159 \\ 
		\hline
		quine-1 & 1 &  71 &  44 & 69 & - & - \\ 
		& 2 & 127 & 142 & 95 & - & - \\ 
		& 3 & 114 & 114 & 93 & - & - \\ 
		\hline
		quine-2 & 1 & 147 & 112 &  56 & - & - \\ 
		& 2 & 161 & 123 & 101 & - & - \\ 
		& 3 & 289 & 189 & 104 & - & - \\ 
		\hline 
		'(I love you)-1 &  99 & 56 & 15 & 22 &  74 & 165 \\ 
		& 198 & 53 & 72 & 55 &  47 &  74 \\
		& 297 & 72 & 90 & 44 & 181 & 365 \\ 
		\hline
		'(I love you)-2 &  99 & 242 &  61 & 16 &  66 &  99 \\ 
		& 198 & 445 & 110 & 60 &  42 &  64 \\
		& 297 & 476 & 146 & 49 & 186 & 322 \\ 
		\hline 
	\end{tabular}
	\caption{The results of a quantitative evaluation: running times of 
	benchmarks 
		in milliseconds}
	\label{compare-efficiency}
\end{table}

In this section, we compare the efficiency of search strategies. A concise 
description is in Table~\ref{compare-efficiency}. A hyphen means running out of 
memory. The first two benchmarks are taken from 
\citet{friedman_reasoned_2018}. \texttt{reverso} is from 
\citet{rozplokhas2018improving}. Next two benchmarks 
about quine are modified from a similar test case in \citet{byrd2017unified}. 
The modifications are made 
to circumvent the need for symbolic constraints (e.g. $\neq$, 
\texttt{absent$^o$}). Our version generates de 
Bruijnized expressions and prevent closures getting into list. The two 
benchmarks differ in the \conde{} clause order of their relational 
interpreters. 
The last two 
benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love 
you)}. This benchmark is also inspired by \citet{byrd2017unified}. Again, the 
sibling benchmarks differ in the \conde{} clause order of their relational 
interpreters. The first one 
has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the 
end, while the other has them at the beginning. We conjecture that DFS$_\textrm{i}$ would 
perform badly in the second case because elimination rules complicate the 
problem when running backward. The evaluation supports our conjecture.

In general, only DFS$_\textrm{i}$ and DFS$_\textrm{bi}$ constantly perform well. DFS$_\textrm{f}$ is just as 
efficient in all benchmarks but \texttt{very-recursiveo}. Both BFS have obvious 
overhead in many cases. Among the three variants of DFS (they all have unfair 
\conj{}), DFS$_\textrm{f}$ is most resistant to clause permutation, followd by DFS$_\textrm{bi}$ then 
DFS$_\textrm{i}$. Among the two implementation of BFS, ours constantly performs as well or 
better. Interestingly, every strategies with fair \disj{} suffers in 
\texttt{very-recursiveo} and DFS$_\textrm{f}$ performs well elsewhere. 
Therefore, this 
benchmark might be a special case. Fair \conj{} imposes considerable overhead 
constantly except in \texttt{appendo}. The reason might be that strategies with 
fair \conj{} tend to keep more intermediate answers in the memory.

@@

Edward points out a disjunct complex would be `fair' if it is a full and 
balanced tree \citet{yang2010adventures}.

Silvija et al \citet{seres1999algebra} also describe a breadth-first search 
strategy. We proof their BFS is equivalent to ours. But our code is shorter
and performs better in comparison with a straightforward translation of their 
Haskell code.

@@

We analysis the definitions of fair \disj{} and fair \conj{}, then propose a 
new definition of fair \conj{}. Our definition is orthogonal with completeness.

>>>>

\addeditor{MVC}
\addeditor{LKC}
\addeditor{DPF}

@@

\title{Towards a miniKanren with fair search strategies}
\author{Kuang-Chen Lu}
\affiliation{Indiana University}
\author{Weixi Ma}
\affiliation{Indiana University}
\author{Daniel P. Friedman}
\affiliation{Indiana University}

@@

\newcommand{\BFSopt}[0]{BFS$_\textrm{opt}$}

\newcommand{\BFSeff}[0]{BFS$_\textrm{eff}$}

\newcommand{\BFSser}[0]{BFS$_\textrm{ser}$}
\newtheorem{defn}{Definition}[section]

@@

  We describe fairness levels in disjunction and conjunction
  implementations.  Specifically, a disjunction implementation can be
  fair, almost-fair, or unfair. And a conjunction implementation can
  be fair or unfair.  We compare the fairness level of four search
  strategies: the standard miniKanren interleaving depth-first search (\DFSi),
  the balanced interleaving depth-first search (\DFSbi), the fair depth-first
  search (\DFSf), and the standard breadth-first search (\BFSser).  
  \DFSbi{} and \DFSf{} are new. And we present a new implementation of 
  \BFSser. Using a quantitative evaluation, we argue that
  \DFSbi{} and \DFSf{} are competitive alternatives to \DFSi, 
  and that \BFSopt{} is more efficient than the standard \BFSser.

@@

miniKanren is a family of relational programming languages.
\citet{friedman_reasoned_2018} introduce miniKanren and its
implementation in \emph{The Reasoned Schemer, 2nd Ed}
(TRS2). miniKanren programs, have been proven to be useful in solving 
many problems as in miniKanren.org.\citet{byrd2017unified}.

% miniKanren is a family of relational programming languages.
% \citet{friedman_reasoned_2018} introduce miniKanren and its implementation
% in \emph{The Reasoned Schemer, 2nd Ed} (TRS2). miniKanren programs, especially
% relational interpreters, have been proven to be useful in solving many problems 
% as in \citet{byrd2017unified}.

\NOTE[DPF]{I think that the abbreviations for the search strategies
should not come up until the Introduction.}

\NOTE[DPF]{When we talk about complete and incomplete, I think we should
make it explicit that we are only talking about (run n (x y z) g ...) and
not (run* (x y z) g ...), but an alternative could be something like is
a stream that is always growing.  There is a term for this, likes makes
progress, but my mind is hiding the word from me.  It will come to me, but
not at this moment.}

\NOTE[LKC]{Shall we delete the sentence about relational interpreters? It is 
misleading: it is emphasizing relational interpreters, but we say nothing 
further about them later.}
\NOTE[DPF]{I have deleted the sentence.}
% \conde{} is the flagship disjunction operator in miniKanren.
A subtlety arises 
when a \conde{} contains many clauses: not every clause has an 
equal chance to contribute to the result. As an example, consider the following 
relation \texttt{repeato} and its invocation.

@@

The miniKanren in TR2 implements interleaving DFS (\DFSi), the cause of this 
unexpected result. With this search strategy, each clause takes half 
of its received computational resources and pass the other half to its 
following clauses, except for the last clause that takes all resources it 
receives. In the example above, the \texttt{a} clause takes half of all 
resourses. And the \texttt{b} clause takes a quarter. Thus \texttt{c} and 
\texttt{d} barely contribute to the result.

@@

\DFSi{} is sometimes powerful for an expert. By carefully organizing the order 
of \conde{} clauses, a miniKanren program can explore more ``interesting'' 
clauses than those uninteresting ones, and thus use computational resources 
efficiently. A little miniKanrener, however, may beg to differ--understanding
implementation details and fiddling with clause order is not the first
priority of a beginner.

@@

The specification that every clause in the same \conde{} is given equal 
``search priority'' is called fair \disj{}. And search strategies with 
almost-fair \disj{} give every clause similar priority. 
Fair \conj{}, a related concept, is more subtle. We defer it to the next 
section.

To summarize our contribution, we
\begin{itemize}
	\item propose and implement balanced interleaving depth-first search 
	(\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement fair depth-first search (\DFSf{}), a 
	new search strategy with fair \disj{}.
	\item implement in a new way breath-first search (\BFSopt), 
        a search strategy 
	with fair \disj{} and fair \conj{}. And we prove formally that our 
\BFSopt{} implementation is semantically equivalent to the one 
by \citet{seres1999algebra}, however, Our code 
runs faster in all benchmarks and we believe it is simpler.
\end{itemize}

@@

Under the hood, the \conde{} here is allocating computational resources to 
four trivially different search space. The unfair \disj{} in 
\DFSi{} allocates many more resources to the first search space. On the 
contrary, fair \disj{} would allocate resources evenly to each search space.

@@

We end this subsection with precise definitions of all levels of 
\disj{} fairness. Our definition of \emph{fair} \disj{} is slightly generalize 
from the one by \citet{seres1999algebra}. Their definition is for binary 
disjunction. We generalize it to a multi-arity one.

\begin{defn}[fair \disj{}]
A \disj{} is fair if and only if it allocates computational resources evenly to 
search spaces produced by goals in the same disjunction (i.e. clauses in 
the same \conde).
\end{defn}

\begin{defn}[almost-fair \disj{}]
A \disj{} is almost-fair if and only if it allocates computational resources
evenly to search spaces produced by goals in the same disjunction that 
the maximal ratio of resources is bounded by a constant.
\end{defn}

@@

A more interesting situation is when the first conjunct produces an unbounded
number of states. Consider the following example, a naive specification of fair \conj{} 
might require search strategies to produce all sorts of singleton lists, but no 
longer ones, which makes the strategies incomplete. A search strategy is 
\emph{complete} if and only if it can determine all the answers within finite 
time, otherwise, it is \emph{incomplete}.

@@

\BFSser{} puts states of the same cost in the same bag, and allocates
resources carefully among search spaces related to different bags such
that answers are produced in increasing order of cost. The \emph{cost}
of an answer is its depth in the search tree (i.e., the number of
calls to relations required to find them) \citep{seres1999algebra}. In
the following example, every answer is a list of list of symbols,
where inner lists are the same. Here the cost of each answer is equal
to the length of its inner lists plus the length of its outer list.

@@

\begin{defn}[fair \conj{}]
A \conj{} is fair if and only if it allocates computational resources evenly to 
search spaces produced from states in the same bag. A bag is a finite 
collection of states. And search strategies with fair \conj{} should represent 
search spaces with possibly unbounded collections of bags. 
\end{defn}

@@

\NOTE[DPF]{If we use the name subscript, the subscripts should use an underscore and we should do it for gunderscorel and gundescorer. If we can use an l that
looks less like a 1, that would even be better.  Look at the l in TLT.}

Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: variables bound to 
states are named \texttt{s}; variables bound to goals are named \texttt{g} 
(possibly with subscript); variables bound to search spaces have names ending 
with $^\infty$. Fig.~\ref{DFSi-0} shows the implementation of \disj. The 
first function, \texttt{disj2}, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \texttt{conj2}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The latter process is done with a 
the helper function \texttt{append-map$^\infty$} that applies its input goal to states 
in its input search spaces and composes the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

Balanced interleaving DFS (DFS$_\textrm{bi}$) has an almost-fair \disj{} and unfair 
\conj{}. The implementation of DFS$_\textrm{bi}$ differs from 
DFS$_\textrm{i}$'s in the \disj{} macro. We list the new \disj{} with its 
helper in Fig.~\ref{balanced-disj}. When there are one or more disjunctive 
goals, \disj{} builds a balanced binary tree whose leaves are the goals and 
whose nodes are \texttt{disj2}s, hence the name of this search strategy. 
The new helper, \texttt{disj+}, takes two additional `arguments'. They 
accumulate goals to be put in the left and right subtrees. The first clause 
handles the case where there is only one goal. In this case, the tree is the 
goal itself. When there are more goals, we partition the list of goals 
into two sublists of roughly equal lengths and recur on the two sublists. Goals 
are moved to the accumulators in the last clause. As we are moving 
two goals each time, there are two base cases: (1) no goal remains; (2) one 
goal remains. These two new base cases are handled in the second clause and the 
third clause, respectively. In contrast, the \disj{} in \DFSi{} constructs the 
binary tree in a particularly unbalanced form.

@@

\NOTE[DPF]{Why does the pair? case precede the null? case? Is it simply more efficient? It shouldn't be?}

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\texttt{disj2} (Fig.~\ref{fDFS}). \texttt{disj2} is changed to call a new and 
fair version of \texttt{append$^\infty$}. \texttt{append$^\infty_{fair}$} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, set to 
\texttt{\#{}t}, which indicates that 
\texttt{s$^\infty$} and \texttt{t$^\infty$} haven't been swapped. The swapping 
happens at 
the third \texttt{cond} clause in the helper, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \texttt{s$^\infty$} and 
\texttt{t$^\infty$} are ended with a thunk and they have been swapped. In this 
case, a new thunk is constructed. Here changing the order of 
\texttt{t$^\infty$} and \texttt{s$^\infty$} won't hurt the fairness (though it 
changes the order of answers). We swap them back so that answers are 
produced in a more natural order.

\NOTE[DPF]{When or where do you swap them back?}

@@

\BFSser{} has both fair \disj{} and fair \conj{}. Our implementation is based on 
DFS$_\textrm{f}$ (not DFS$_\textrm{i}$). To implement \BFSopt{} based on \DFSf{}, 
we need \texttt{append-map$^\infty_{fair}$} in addition to 
\texttt{append$^\infty_{fair}$}. 
The only difference between \texttt{append-map$^\infty$} and 
\texttt{append-map$^\infty_{fair}$} is that the latter calls 
\texttt{append$^\infty_{fair}$} instead of \texttt{append$^\infty$}.

The implementation can be improved in two ways. First, as mentioned in 
section 2.2, \BFSser puts answers in bags and answers of the same cost are in the 
same bag. In this implementation, however, it is unclear where this information 
is recorded. Second, \texttt{append$^\infty_{fair}$} is extravagant in memory 
usage. It 
makes $O(n+m)$ new \texttt{cons} cells every time, where $n$ and $m$ are the 
``length''s of input search spaces. We address these issues in the first 
subsection.

Both \BFSser{} \citet{seres1999algebra} and BFSopt{} produce answers in 
increasing order of cost. So it is interesting to see if they are equivalent. 
We prove so in Coq. The details are in the second subsection.

@@

\NOTE[DPF]{Never refer to a literal section like 2.2. Instead
stick a label where it starts right near the section number and
stick another label where the subsection starts.
Then ref the two labels and place a period between them.}.

As mentioned in section 2.2, BFS puts answers in bags and answers of the 
same cost are in the same bag. The cost
information is recorded subtly -- the \texttt{car}s of a search space have cost 
0 (i.e. they are all in the same bag), and the costs of answers in thunk are 
computed recursively then increased by one. It is even more subtle that
\texttt{append$^\infty_{fair}$} and the \texttt{append-map$^\infty_{fair}$} 
respects the cost information. We make these facts more obvious by 
changing the type of search space, modifying related function definitions, 
and introducing a few more functions.

@@

Functions related to the pure subset are listed in Fig.~\ref{BFS-opt} (the 
others in Fig.~\ref{BFS-opt-cont}). They are compared with 
\citeauthor{seres1999algebra}'s implementation in our proof. The first three 
functions in Fig.~\ref{BFS-opt} are search space constructors. \texttt{none} 
makes an empty search space; \texttt{unit} makes a space from one answer; and 
\texttt{step} makes a space from a thunk. The remaining functions as before.

Luckily, the change in \texttt{append$^\infty_{fair}$} also fixes the miserable 
space 
extravagance---the use of \texttt{append} helps us to reuse the first bag of 
\texttt{t$^\infty$}.

\citet{kiselyov2005backtracking} has shown that a \emph{MonadPlus} hides in 
implementations of logic programming system. The \BFSopt{} implementation is not 
an exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, 
but takes arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, 
and \texttt{mplus} respectively.

@@

\subsection{comparison with the BFSser and BFSopt}

In this section, we compare the pure subset of \BFSopt{} with the BFS 
found in \citet{seres1999algebra}. We focus on the pure subset because 
their system is pure. Their system represents search spaces with streams of 
lists of answers, where each list is a bag.

To compare efficiency, we translate their Haskell code into Racket (See 
supplements for the translated code). The translation is direct 
due to the similarity in both logic programming systems and search space 
representations. The translated code is longer and slower than BFSopt
implementation. Details about 
difference in efficiency are in section 7.

We prove in Coq that \BFSser{} and \BFSopt{} are semantically equivalent, i.e. \texttt{(run n ? g)} 
produces the same result (See supplements for the formal proof).

@@

\begin{table}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline 
		benchmark & size & DFS$_\textrm{i}$ & DFS$_\textrm{bi}$ & DFS$_\textrm{f}$ & BFS$_\textrm{opt}$ & BFS$_\textrm{ser}$  
		\\
		\hline
		very-recursiveo & 100000 &  579 &  793 & 2131 & 1438 & 3617 \\
		& 200000 & 1283 & 1610 & 3602 & 2803 & 4212 \\
		& 300000 & 2160 & 2836 &    - & 6137 &    - \\
		\hline 
		appendo  & 100 &  31 &  41 &  42 &  31 &  68 \\ 
		& 200 & 224 & 222 & 221 & 226 & 218 \\ 
		& 300 & 617 & 634 & 593 & 631 & 622 \\ 
		\hline 
		reverso & 10 &   5 &   3 &   3 &     38 &     85 \\ 
		& 20 & 107 &  98 &  51 &   4862 &   5844 \\
		& 30 & 446 & 442 & 485 & 123288 & 132159 \\ 
		\hline
		quine-1 & 1 &  71 &  44 & 69 & - & - \\ 
		& 2 & 127 & 142 & 95 & - & - \\ 
		& 3 & 114 & 114 & 93 & - & - \\ 
		\hline
		quine-2 & 1 & 147 & 112 &  56 & - & - \\ 
		& 2 & 161 & 123 & 101 & - & - \\ 
		& 3 & 289 & 189 & 104 & - & - \\ 
		\hline 
		'(I love you)-1 &  99 & 56 & 15 & 22 &  74 & 165 \\ 
		& 198 & 53 & 72 & 55 &  47 &  74 \\
		& 297 & 72 & 90 & 44 & 181 & 365 \\ 
		\hline
		'(I love you)-2 &  99 & 242 &  61 & 16 &  66 &  99 \\ 
		& 198 & 445 & 110 & 60 &  42 &  64 \\
		& 297 & 476 & 146 & 49 & 186 & 322 \\ 
		\hline 
	\end{tabular}
	\caption{The results of a quantitative evaluation: running times of 
	benchmarks 
		in milliseconds}
	\label{compare-efficiency}
\end{table}

Here we compare the efficiency of search strategies. A concise 
description is in Table~\ref{compare-efficiency}. A hyphen means ``running out of 
memory.'' The first two benchmarks are taken from 
TRS2. \texttt{reverso} is from 
\citet{rozplokhas2018improving}. The next two benchmarks 
about quine are modified from a similar test case in \citet{byrd2017unified}. 
The modifications are made 
to circumvent the need for symbolic constraints (e.g. $\neq$, 
\texttt{absent$^o$}). Our version generates de 
Bruijnized expressions and prevent closures getting into list. The two 
benchmarks differ in the \conde{} clause order of their relational 
interpreters. 
The last two 
benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love 
you)}. This benchmark is from \citet{byrd2017unified}. Again, the 
sibling benchmarks differ in the \conde{} clause order of their relational 
interpreters. The first one 
has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the 
end, while the other has them at the beginning. We conjecture that DFS$_\textrm{i}$ would 
perform badly in the second case because elimination rules complicate the 
problem when synthesizing (i.e., our evaluation supports our conjecture.)

In general, only DFS$_\textrm{i}$ and DFS$_\textrm{bi}$ constantly perform well. DFS$_\textrm{f}$ is just as 
efficient in all benchmarks but \texttt{very-recursiveo}. Both BFS have obvious 
overhead in many cases. Among the three variants of DFS (they all have unfair 
\conj{}), DFS$_\textrm{f}$ is most resistant to clause permutation, followed by DFS$_\textrm{bi}$ then 
DFS$_\textrm{i}$. Among the two implementation of BFS, ours constantly performs as well or 
better. Interestingly, every strategy with fair \disj{} suffers in 
\texttt{very-recursiveo} and DFS$_\textrm{f}$ performs well elsewhere. 
Therefore, this 
benchmark might be a special case. Fair \conj{} imposes considerable overhead 
constantly except in \texttt{appendo}. The reason might be that strategies with 
fair \conj{} tend to keep more intermediate answers in the memory.

@@

Yang \citet{yang2010adventures} points out a disjunct complex would be `fair' 
if it were a full and balanced tree .

Seres et al \citet{seres1999algebra} also describe a breadth-first search 
strategy. We prove \BFSser{} is semantically equivalent to \BFSopt. But our codeis shorter and performs better in comparison with a straightforward 
translation of their Haskell code.

@@

We analyze the definitions of fair \disj{} and fair \conj{}, then propose a 
new definition of fair \conj{}. Our definition is orthogonal with completeness.

We devise two new search strategies (i.e. balanced interleaving DFS 
(DFS$_\textrm{bi}$) and fair DFS (DFS$_\textrm{f}$)) and devise a new 
implementation of BFS. These strategies have different features 
in fairness: $_\textrm{bi}$ has an almost-fair \disj{} and unfair \conj{}. 
DFS$_\textrm{f}$ has fair \disj{} and unfair \conj{}. \BFSopt{} has both fair
\disj{} and fair \conj{}.

Our quantitative evaluation shows that DFS$_\textrm{bi}$ and DFS$_\textrm{f}$ are competitive 
alternatives to DFS$_\textrm{i}$, the current search strategy, and that 
\BFSopt is less practical than 
\NOTE[DPF]{what}
